### 2016-08-02 main.tex
Add a reference that Sam Sober mentioned in his review--a nice application.

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

% Remove comment for double spacing
\usepackage{setspace} 
\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

%%% Begin BWP
%\usepackage{amsmath, amsthm, amssymb, wasysym, graphicx}
%\usepackage[small, hang, bf]{caption}
%\usepackage{natbib}
%\renewcommand\cite{\citep}
%\newcommand\citepossessive[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand\eq[1]{Eq.~\ref{#1}}
\newcommand\fig[1]{Fig.~\ref{#1}}
\newcommand\sref[1]{Section~\ref{#1}}
\let\oldmarginpar\marginpar
\renewcommand{\marginpar}[1]{\oldmarginpar{\linespread{1}\scriptsize{#1}}}

% PLOS wants \paragraph for some reason...
\renewcommand{\subsubsection}[1]{\paragraph{#1}}

\setlength{\marginparwidth}{55mm}

\newcommand\argmin{\mathop{\textrm{{\rm argmin}}}\limits}
\newcommand{\noprint}[1]{}

\hyphenation{cross-valid-ation}
%%% End BWP

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{A Fast and Accurate Zebra Finch Syllable Detector}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Ben Pearre\textsuperscript{1,a,\textcurrency},
L.~Nathan Perkins\textsuperscript{1,b},
Jeffrey E.~Markowitz\textsuperscript{2,c},
Timothy J.~Gardner\textsuperscript{1,d}
\\
\bigskip
\textsuperscript{1} Department of Biology, Boston University, Boston, Massachusetts, United States of America\\
\textsuperscript{2} Department of Neurobiology, Harvard Medical School, Boston, Massachusetts, United States of America
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.
Author \textsuperscript{a}~contributed the original MATLAB and LabVIEW implementations and most of the manuscript.
\textsuperscript{b}~contributed the Swift implementation and its description herein, helped debug the training code, and measured the timing for most of the detectors.
\textsuperscript{c}~performed the fibre photometry experiment, wrote several types of filter-bank software against which we compared, made suggestions for the neural network detector, and reformatted the MATLAB code for distribution.
\textsuperscript{d}~provided a stream of suggested techniques, metrics, and goals.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency a Insert current address of first author with an address update
% \textcurrency b Insert current address of second author with an address update
% \textcurrency c Insert current address of third author with an address update

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
\textsuperscript{\textcurrency} Corresponding author: bwpearre@gmail.com (BP)

\end{flushleft}

\reversemarginpar
%%% End PLOS header template

\begin{abstract}
  The song of the adult male zebra finch is strikingly stereotyped.
  Efforts to understand motor output, pattern generation, and learning
  have taken advantage of this consistency by investigating the bird's
  ability to modify specific parts of song under external cues, and by
  examining timing relationships between neural activity and vocal
  output.  Such experiments require that precise moments during song
  be identified in real time as the bird sings.  Various
  syllable-detection methods exist, but many require special hardware,
  software, and know-how, and details on their implementation and
  performance are scarce.  We present an accurate, versatile, and fast
  syllable detector that can control hardware at precisely timed
  moments during zebra finch song.  Many moments during song can be
  isolated and detected with false negative and false positive rates
  well under 0.5\% and 0.01\% respectively. The detector can run on a
  stock Mac Mini with triggering delay of less than a millisecond and
  a jitter of $\sigma\approx 2$ milliseconds.
\end{abstract}

\linenumbers

\section{Introduction}

The adult zebra finch ({\em Taeniopygia guttata}) sings a song made up
of 2--6 syllables, with longer songs taking on the order of a
second. The song may be repeated hundreds of times per day, is almost
identical each time, and several brain areas reflect this consistency
in highly stereotyped neural firing patterns. This consistency makes
the zebra finch one of the most popular models for the study of the
neural basis of learning, audition, and control.

This consistency allows a variety of experiments if precise moments in
song can reliably be detected quickly enough to trigger other
apparatus during singing.  A common area of study with this
song-triggering technique is the anterior forebrain pathway (AFP), a
homologue of mammalian basal ganglia consisting of a few distinct
brain areas concerned with the acquisition and learning of song.  For
example, \cite{Kao2005} stimulated the lateral magnocellular nucleus
of the anterior nidopallium (LMAN)---the output nucleus of the
AFP---at precisely timed moments during song and showed that this area
repeatably controls specific variables in song output.
\cite{Wang2008interhemispheric} stimulated LMAN and the high vocal
centre (HVC) in one hemisphere or the other, and found that control of
song rapidly switches between hemispheres.  \cite{Keller2009}
investigated the response of cells in field L and the caudolateral
mesopallium during song with and without feedback perturbations as
well as external sounds, and concluded that these regions may hold a
representation of song against which auditory signals are compared.
\cite{Tumer2007pitchshift} played a disruptive sound during renditions
of a syllable that were slightly above (or below) its average pitch,
and showed that the presence of this apparently random natural
variability in songbird motor output is used to drive change in the
song.  \cite{Andalman2009} extended this experiment and showed that
AFP produces a corrective signal to bias song away from those
disruptions.  \cite{Charlesworth2011learning} showed that the song
change was isolated to within roughly 10 milliseconds (ms) of the
stimulus, and further proposed an elegant model that predicts the
shape of the learned response and suggests a simple mechanism.
\cite{Warren2011} showed that that AFP transfers the error signal to
the robust nucleus of the arcopallium (RA) using
NMDA-receptor--mediated glutamatergic transmission. By looking at song
recovery after applying the same pitch-shift paradigm,
\cite{Canopoli2014} showed that the caudal medial nidopallium is
implicated in memorising or recalling a recent song target, but for
neither auditory processing nor directed motor learning.

% THEIR WORDS: (NCM, a high auditory area) impaired recovery of the original pitch even several weeks after withdrawal of the reinforcing stimuli. Because NCM lesions spared both successful noise-avoidance behavior and birds' auditory discrimination ability, our results show that NCM is not needed for directed motor changes or for auditory discriminative processing, but is implied in memorizing or recalling the memory of the recent song target.

Despite the power and versatility of such experiments, there is no standard syllable detector.  Considerations include:
\begin{description}
\item[Accuracy:] How often does the system produce false positives or false negatives?
\item[Latency:] The average delay between the target syllable being sung and the detection.
\item[Jitter:] The amount that latency changes from instance to instance of song.  Our measure of jitter is the standard deviation $\sigma$ of latency.
\item[Versatility:] Is detection possible at ``difficult'' syllables?
\item[Ease of use:] How automated is the process of programming a detector?
\item[Cost:] What are the hardware and software requirements?
\end{description}

A variety of syllable-triggering systems have been used, but few have
been documented or characterised in detail.  In 1999,
\cite{Leonardo1999} used groups of IIR filters with hand-tuned logical
operators.  Their system had a latency of 50 or 100 ms,
and they do not report on jitter or accuracy.  As access to
computational resources has improved, approaches have changed: in
2009, \cite{Andalman2009} still used hand-tuned filters, but ran them
on a Tucker-Davis Technologies digital signal processor.  They report
a latency of around 4 ms, but as with other filter-bank techniques, it
is not strictly a syllable detector but rather a pitch and timbre
detector---it cannot identify a frequency sweep, or distinguish a
short chirp from a long one---and thus requires careful selection of
target syllables.  Furthermore, the method is neither inexpensive nor,
based on our experience with a similar technique, accurate.
\cite{Keller2009} applied a neural network to a spectral image of
song.  They report a jitter of 4.3 ms, but further implementation and
performance details are not available.  In 2011, \cite{Warren2011}
matched spectral templates to stable portions of syllables in 8-ms
segments.  They report a jitter of 4.5 ms, and false-negative and
false-positive rates of up to 2\% and 4\%, respectively.  Hardware
requirements and ease of use were not reported.  In 2013,
\cite{Skocik2013} described in detail a system that matches spectral
images of template syllables using a correlation coefficient.  With a
fast desktop (Intel i7 six-core) running Linux and equipped with a
National Instruments data acquisition card, it boasts a hardware-only
(not accounting for the time taken to compute a match with a syllable)
latency and jitter of just a few microseconds, and the detection
computation they use should not much increase that.  Drawbacks are
that some hand-tuning may still be required, and that they report
false-negative rates around 4\% and 7\% for zebra finches and
Bengalese finches respectively, measured on a small dataset.  In much
other work, an allusion is made to a syllable detector, but no further
information is provided.

We developed a standalone detector that learns to match moments in the
song using a neural network applied to the song spectrogram, and
outputs a TTL pulse (a brief 5-volt pulse) at the chosen moment. The approach consists of
three steps:

\begin{enumerate}
\item Record and align a corpus of training songs.  The technique has been published in \cite{Poole2012}.
\item Choose one or more instants in the song that should create a trigger event, and train a
  neural network to recognise them. This step is carried out offline.  While any neural network software would produce equivalent results, we used MATLAB 2015b's neural network toolbox.
\item Once trained and saved, the neural network is used by a realtime
  detection program that listens to an audio signal and indicates detection of the target
  syllables via a TTL pulse.  We present three implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
\end{enumerate}
This method makes the following contributions:
\begin{itemize}
\item Fast: sub-millisecond latencies, with jitter around 2 ms.
\item Accurate: false positive rates under 0.02\% and false negative rates under 0.5\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
\item State-of-the-art performance with default parameters.
\item Runs on inexpensive hardware.
\item Described in detail here, with reference implementations provided and benchmarked.
\end{itemize}

%We present the method in Section~\ref{sec:method}.  Section~\ref{sec:results} presents our measurements.
%We
%conclude in Section~\ref{sec:conclusion}, and point to software
%resources in Appendix~\ref{sec:resources}.

\section{Materials and Methods}
\label{sec:method}

\subsection{Learning a detector}

We begin with a few hundred recordings of a given bird's song, as well as calls, cage noise, and other non-song audio data.  Male
zebra finch song is most highly stereotyped when a female is present
(``directed song'') and slightly more variable otherwise
(``undirected''); we train and test on undirected song since this is
more challenging.  Recordings were made inside a sound-isolating
chamber in which was mounted a recording microphone (Audio-Technica
AT803).  The songs were time-aligned as described in \cite{Poole2012}.

We rely on two circular buffers:
\begin{description}
\item[Audio buffer:] This contains the most recent audio samples, and
  is of the length required to compute the Fast Fourier Transform
  (FFT)---usually 256 samples.
\item[FFT buffer:] The results of each new FFT are placed here.  It
  contains the $n_\textrm{fft}$ most recent FFTs, which will serve as
  inputs to the neural network (described below).
\end{description}

Audio is recorded at some sample rate $1/t_\mathrm{sample}$ (for example, 44.1 kHz), and new data are appended to the circular audio buffer.

The spectrogram is computed at regular intervals---the useful range
seems to be roughly 1--5 milliseconds, which we refer to as the frame
interval $t_\textrm{fft}$.  At each frame, a spectrum is computed from
the most recent 256 audio samples in the audio buffer, and the result
is appended to the FFT buffer.  For example, if $t_\textrm{fft}=1$ ms
and the recording sample rate $1/t_\textrm{sample}=40$ kHz, then in
order to compute a new FFT, $40000\cdot 0.001=40$ new audio samples
must be appended to the audio buffer, the FFT is computed using the
most recent 256 samples in that buffer, and the result is appended to
the FFT buffer.  Over time, the successive spectra will look something
like \fig{fig:song}.

\begin{figure}
  \includegraphics[width=\textwidth]{Fig1}
  \caption{This image was made by superposing the spectra of our 2818
    aligned songs.  Our example detection points, $t^*_1\ldots t^*_6$,
    are shown as red lines, with the recognition regions (30 ms
    $\times$ 1--8 kHz) marked as rectangles.}
  \label{fig:song}
\end{figure}

Note that this results in time being discretised into chunks of length
$t_\mathrm{fft}$.  Because each Fourier transform computation contains
a small number $n_\mathrm{fft}$ of new audio samples (in the example
above, 40 new samples and $256-40$ that have already been examined),
we tried implementing the Sliding Discrete Fourier transform (SDFT)
\cite{Jacobsen2003SlidingDFT}.  This allows
$t_\textrm{fft}=t_\textrm{sample}$.  Practically, the operating system
retrieves new audio samples from the hardware several at a time, so
the full benefit of the SDFT is difficult to see in practice.
Furthermore, we found that FFT implementations are sufficiently highly
optimised that discretising time into chunks of $t_\mathrm{fft}$ as we
have done produced similar results with simpler software.

When using consumer audio hardware that operates best at 44.1 kHz, the
requested frame interval may not line up with the sample rate, so the
actual frame interval may be different from the intended.  For
example, at 44.1 kHz a 1-ms frame interval requires a new FFT every
44.1 samples.  This must be rounded to 44 samples, resulting in
$t_\textrm{fft}=\lfloor 44.1 \rceil / 44.1 \approx 0.9977$ ms.

One or more target moments during the song must be chosen.  Our
interface presents the time-aligned spectrogram averaged over training
songs, and requires manual input of the target times, $t^*$.  Then we
assemble the training set from the song data, train the network,
compute optimal output unit thresholds, and save the network object
and an audio test file.

\subsubsection{Recognition region}
\label{sec:recognitionregion}

The neural network's inputs are the FFT values from a rectangular
region of the spectrogram covering a predefined range of frequency
values $F$ (for example, 1--8 kHz) at some number of the most recent
frames $n_\textrm{fft}$.  Any time $t$ falls within frame
$\tau(t)$, and $t-t_\textrm{fft}$ falls within the previous frame, so the recognition region that the neural network receives as input
consists of the spectrogram values over $F$ at $\tau(t)$ and those
from the contiguous set of recent frames: $T = \{$ $\tau(t)$,
$\tau(t-t_\textrm{fft})$, $\tau(t-{2t_\textrm{fft}})\ldots$
$\tau(t-n_\textrm{fft}t_\textrm{fft})\}$.  We have found that a 30-ms
time window---which will yield $n_\textrm{fft} = |T|=\lfloor 30\textrm{
  ms}/t_\textrm{fft}\rceil$ frames---of frequencies spanning 1--8 kHz
works well for most syllables.

Six examples of chosen target moments in the song, and their
corresponding recognition regions $F$ and $T$, are shown in
\fig{fig:song}.

%% \subsubsection{Song micro-realignment}

%% As conditions change, and especially during undirected song, syllable
%% length and relative timing may vary slightly, which introduces
%% variations in the precise timing of each syllable. The alignment
%% software we use ensures that songs are aligned at the point midway
%% through the song, but if the target syllable is not at that point, it
%% is helpful to re-align the songs at the point of interest.  This may
%% be accomplished by looking for peaks in the correlation of the
%% time-domain signal with the song whose spectrogram is closest to
%% average over the training set.\marginpar{I hope that's the right thing
%%   to do. Seems to work, anyway\dots}

\subsubsection{Building the training set}

The training set is created in a fashion typical for neural networks:
at each time frame $t$ the rectangular $|F|\times |T|$ recognition
region in the spectrogram as of time $t$ is reshaped into a vector
$\xi_t$, which will have length $|F||T|$ and contain the spectra in
$F$ taken at all of the times in the set $T$: from
$\tau(t-n_\textrm{fft}t_\textrm{fft})$ to $\tau(t)$.  These vectors
are placed into a training matrix, $\Xi$, such that each column
$\xi_t$ holds the contents of the recognition region---containing
multiple frames from the spectrogram---as of one value of $t$.

Training targets $y_t$ are vectors with one element for each desired
detection syllable.  That element is, roughly, 1 if the input vector
comes from the target time ($t=t^*$), 0 otherwise, for each target
syllable (of which there may be any number, although they increase
training time, and in our implementations the number of distinct
output pulses is constrained by hardware).  Since the song alignment
may not be perfect, and due to sample aliasing, a
strict binary target may ask the network to learn that, of two
practically identical frames, one should be a match and the other
not. Thus it is preferable to spread the target in time, such that at
the target moment it is 1, and at neighbouring moments it is
nonzero. We found that a Gaussian smoothing kernel around the target
time with a standard deviation of 2 ms serves well.

With inputs well outside the space on which a neural network has been
trained, its outputs will be essentially random. In order to reduce
the false positive rate it is necessary to provide negative training
examples that include silence, cage noise, wing flapping, non-song
vocalisations, and perhaps songs from other birds.  Although it will
depend on the makeup of the non-song data, we have found that training
with as low as a 1:1 ratio of non-song to song yields excellent
results on most birds.  These data are also included in our accuracy tests.

\subsubsection{Normalisation}

In order to present consistent and meaningful inputs to the neural
network, we normalise the incoming data stream so that changes in the
song structure of the sound are emphasised over changes in volume, and
to maximise the effectiveness of the neural network's training
algorithm.

The first normalisation step is designed to eliminate differences in
amplitude due to changes in the bird's location and other variations
in recording.  Each recognition region vector $\xi_t$---during
training, each {\em column} of the training matrix $\Xi$---is
normalised using MATLAB's {\tt zscore} function $\hat{\xi}_t = (\xi_t
- \mu_{\xi_t}) / \sigma_{\xi_t}$, so that the content of each window
has mean $\mu_{\hat{\xi}_t}=0$ and standard deviation
$\sigma_{\hat{\xi}_t}=1$.

The second step is designed to ensure that the neural network's inputs
have a range and distribution for which the training function can
easily converge.  Each element $i$ of $\hat{\xi}$ is normalised across
the entire training set---each {\em row} of $\Xi$---in the same way:
$\check{\xi}^i = (\hat{\xi}^i -
\mu_{\hat{\xi}^i})/\sigma_{\hat{\xi}^i}$, so that the values of that
point across the whole training set have mean $\mu_{\check{\xi}^i}=0$
and standard deviation $\sigma_{\check{\xi}^i}=1$.  This is
accomplished during training by setting MATLAB's neural network
toolbox normalisation scheme to {\tt mapstd}, and the scaling
transform is saved as part of the network object used by the realtime
detector so that it may be applied to unseen data at runtime.

These two normalisation steps provide a set of inputs that are more
robust to outliers and less likely to produce false positives during
silence than other normalisation schemes, such as linear or
L\textsubscript{2} normalisation.

\subsubsection{Neural Networks}

While any learned classifier might suffice, we chose a two-layer
feedforward neural network.  In brief, our network takes an input
vector $\xi_t$---as described above---and produces an output vector
$y_t$, and when any element of $y_t$ is above a threshold (described
below), the detector reports a detection event.  The network uses two
matrices of weights, $W_0$ and $W_1$, and two vectors of biases, $b_0$
and $b_1$.  The first takes the input $\xi_t$ to an intermediate
stage---the ``hidden layer'' vector.  To each element of this vector
is applied a nonlinear squashing function such as {\tt tanh}, and then
the second weight matrix $W_1$ is applied.  This produces output
$y_t$:
\begin{equation*}
  y_t = W_1 \tanh (W_0 \xi_t + b_0) + b_1
\end{equation*}
During the network's training phase, the elements of the matrices and bias vectors are learned by back-propagation of errors
over a training set.  A more detailed explanation of neural networks
may be found in \cite{hkp}.

Essentially, after training, the network is available in the form of
the two matrices $W_0$ and $W_1$ and the two vectors $b_0$ and $b_1$,
and running the network consists of two matrix multiplications, two
vector additions and the application of the squashing function.

\subsubsection{Training the network}

The network is trained using MATLAB's neural network toolbox, with Scaled Conjugate Gradient ({\tt trainscg}). We tried
a variety of feedforward neural network geometries, from simple
1-layer perceptrons to geometries with many hidden nodes, as well as
autoencoders. Perhaps surprisingly, even the former yields excellent
results on many syllables, but a 2-layer perceptron with a very small
hidden layer---with a unit count 2-4 times the number of target
syllables---was a good compromise between accuracy and training
speed. For more variable songs, deep structure-preserving networks may
be more appropriate, but they are slow to train and unnecessary for
zebra finch song.

\subsubsection{Computing optimal output thresholds}
\label{sec:optimalthresholds}
After the network is trained, outputs of the network for any input are
now available, and will be in (or, due to noisy inputs and imperfect
training, close to) the interval (0, 1). We must choose a threshold
above which the output is considered a positive detection. Finding the
optimal threshold requires two choices. The first is the relative cost
of false negatives to false positives, $C_n$. The second is the
acceptable time interval: if the true event occurs at time $t$, and
the detector triggers at any time $t\pm\Delta t$, then it is
considered a correct detection. Then the optimal detection threshold
is the number that minimises $\textrm{[false positives]}
+C_n\cdot\textrm{[false negatives]}$ over the training set, using the
definitions of false positives and negatives given in
Section~\ref{sec:accuracy}. Since large portions of the cost function
are flat, we use a brute-force linear search, which requires fractions of a
second. For the results presented here, we have used
$\Delta t=10$ ms, and arbitrarily set $C_n=1$.

\subsubsection{De-bouncing}

During runtime, the network may produce above-threshold responses to nearby
frames.  Thus, after the first response, subsequent responses are suppressed
for 100 ms.  However, for the accuracy figures presented here, we used
the un-de-bounced network response.

\subsubsection{Our parameter choices}

We used an FFT of size 256; a Hamming window; and chose a target
spectrogram frame interval of $t_\textrm{fft}=1.5$ milliseconds,
resulting in a true frame interval of $t_\textrm{fft}=\lfloor 1.5\cdot
44.1\rceil /44.1\approx 1.4966$ ms.  We set the network's input space
to 30 ms long, and to span frequencies from 1--8 kHz, which contains
the fundamentals and several overtones of most zebra finch
vocalisations.

We found these parameters to work well across a variety of target
syllables, but various other parameter sets yield results similar to
those presented here.  Some of the parameters trade off detection
accuracy or temporal precision vs.~training time. For example,
decreasing the frame interval generally decreases both latency and
jitter, but also increases training time.

\subsection{Realtime detection}

The architecture of the realtime detector requires that the most
recent $n_\textrm{fft}$ spectrograms be fed to the neural network
every frame interval.  Audio samples from the microphone are appended
to the circular audio buffer.  Every $t_\textrm{fft}$ seconds a new
spectrogram is calculated by applying the Hamming window to the
contents of the buffer, performing an FFT, and extracting the
power. Outputs of the spectrogram from the target frequency band are
appended to the circular FFT buffer.  The spectrograms are sent to a
static implementation of the previously trained neural network.

We tested three implementations of the realtime detector.  For all of
these tests, we ran the detector processes under the operating
systems' default schedulers and process priorities, running typical
operating system daemon processes but no user loads.  The computers
had ample memory resources.

\subsubsection{Swift}

This detector uses the Swift programming language and Core Audio
interface included in Apple's operating systems.

The Core Audio frameworks provide an adjustable hardware buffer size
for reading from and writing to audio hardware (different from our two
circular buffers). Tuning this buffer size provides a tradeoff between
the jitter in the detection and the processor usage needed to run the
detector. We used buffer sizes ranging from 8 samples (0.18 ms at 44.1
kHz) to 32 samples (0.7 ms at 44.1 kHz) depending on the frame size 
used by the detector.

Vector operations---applying the Hamming window, the FFT, input
normalisation, matrix multiplication, and the neural network's
transfer functions---are performed using the Accelerate framework
(vDSP and vecLib), which use modern vector-oriented processor
instructions to perform calculations.

When the neural network detects a match, it instructs the computer to
generate a TTL pulse that can be used to trigger downstream hardware.
This pulse can be either written to the computer's audio output buffer
(again, in 8- to 32-sample chunks) or sent to a microcontroller
(Arduino) via a USB serial interface. Sending the trigger pulse via
the serial interface and microcontroller is noticeably faster (2.2 ms
lower latency), likely due to the fact that the audio buffer goes
through hardware mixing and filtering prior to output.

The above code can be run on multiple channels of audio on consumer
hardware (such as a 2014 Mac Mini) with little impact on CPU usage
($<15\%$). Depending on the experimental needs, latency can potentially
be further decreased (at the expense of processor usage) by adjusting the audio
buffer sizes.

We ran the Swift detector on a Late 2014 Mac Mini with a Intel Core i5
processor at 2.6GHz  with 16 gigabytes of RAM, running Mac OS X 10.11.

% technically, it is LabVIEW
\subsubsection{LabVIEW}

This implementation requires special software and hardware: LabVIEW
from National Instruments---we used 2014 service pack 1---and a data acquisition card---we use the
National Instruments PCI-6251 card on a PC with an Intel Core i5-4590
processor at 3.7GHz (a relatively low-end machine) with 32 gigabytes
of RAM, running Microsoft Windows 8.1 Pro.

This implementation has several drawbacks: it requires Windows and
expensive hardware and software and due to the programming language it
is difficult to modify and debug---indeed, a persistent bug in our
implementation currently renders it substantially less accurate than
the other detector implementations on some syllables. However, our
test configuration usually achieved excellent performance, and further
gains should be possible if the implementation were retargeted onto
field-programmable gate array (FPGA) hardware---which would have the additional benefit of providing
deterministic ``hard realtime'' guarantees---or just run on a faster
desktop system.

\subsubsection{MATLAB}

This detector uses the built-in audio input and output hardware on a
compatible computer.  We tested on a 2014 Mac Mini (the same machine used
for the Swift detector described above) and 2015 Mac Pro.
The Mac Pro does not have an audio input jack, so input was through an
M-Audio MobilePre external USB audio interface.  Despite the faster
processor, the latter system did not achieve higher performance than
the former, due to USB data transfer overhead.

Because of how MATLAB's DSP toolbox interfaces with audio hardware,
there is a double buffer both reading from and writing to audio
hardware. As a result, much of the code focuses on a lightweight audio
hardware interface, in order to have the smallest achievable audio
buffer. To help achieve this, the MATLAB implementation spreads data
acquisition and processing across two threads, due to the higher
computational overhead of the interpreted programming language.

The most versatile implementation, MATLAB runs on a variety of
hardware and operating systems, and is perhaps the easiest to modify.
While it did not perform as well on our test system as the other
implementations, the convenience may outweigh the slight timing
performance penalty for some experiments.  Key to minimising jitter is
the size of the audio buffer: on a 2014 Mac Mini running MATLAB 2015b
the smallest buffer size that did not result in read overruns was
about 4 ms.

Similar to the Swift implementation, it is also possible to modify the
MATLAB implementation to generate the TTL pulse from a microcontroller
(Arduino) controlled via a serial over USB interface. This eliminates
the double buffer required when sending triggers via the audio
interface, reducing latency by half.

\subsection{Quantification}
\label{sec:quantify}

We divided the data into training and test sets.  For the results
reported in the following section, the training set consisted of 1000
instances of the song of one bird, and 2000 song-length samples of
non-song (cage noise, calls, other birds' song).  Our test set
consisted of an additional 1818 iterations of the bird's song and 3636
song-length samples of non-song.

The MATLAB neural network toolbox further divides our ``training'' set
into internal training, validation, and test sets.  Because we do our
own testing on a distinct and much larger data set, we instructed the
toolbox to use 80\% of its training set for training, and the
remaining 20\% for validation, with no hold-out test data.

Because the data set consists of temporally aligned songs, ground
truth is available for each song (albeit with minor variations due to
the alignment process used \cite{Poole2012}), and thus the detector
can be checked by presenting the recorded training songs as well as
the canonical detection events. To this end, besides the trained
network object, our learning code produces an audio file consisting of
all of the training data on the left audio channel and a delta
function at each aligned moment of target syllable presentation on the
right channel. Thus, when played on any audio player, the left channel
may be provided as input to the detector, and the the detector's
output pulses may be compared against the right channel.

\subsubsection{Accuracy}
\label{sec:accuracy}

We define the accuracy of the network based on its classification
performance per frame. In order to avoid the apparent problem of every
non-detected non-syllable counting as a true negative, we also tried
defining accuracy on a per-song basis, such that a song without the
target syllable counted as a single true negative.  Computing the
optimal output thresholds on a per-frame basis resulted in higher
thresholds and thus a lower false-positive rate, with minimal
consequences to the false-negative rate, while also providing a valid
definition of false-positive rate for data streams that had not been
segmented into song-sized chunks and thus allowing easy interpretation
of false-positive rates from calls, cage noise, etc.

%% We define accuracy on a per-song basis, as follows:

%% \begin{description}
%%   \item[True positive:] A song for which the detector fires within
%%     20ms of the intended moment.
%%   \item[True negative:] So that every non-detected non-syllable does
%%     not count, a true negative is a complete song during which the
%%     detector did not fire at any point outside the true positive
%%     region.
%%   \item[False positive:] A detection event more than 20ms from the
%%     target syllable.\marginpar{Or a song with $>0$ false pos?}
%%   \item[False negative:] A song's target interval during which a
%%     detection event should have occurred, but didn't.
%% \end{description}

The accuracy as defined above is used for computing the optimal
thresholds above which the network's output should be interpreted as a
match on the training data as described in
\sref{sec:optimalthresholds}, for evaluation of the detectors on the
training songs, and while live.

\noprint{ Since the network will output values $o_t$ between 0 and 1
  at each moment $t$ in an attempt to match the training output, the
  optimal threshold $\tau\in[0,1]$ for the output neuron should be
  computed.  Given the relative cost of false positives vs.~false
  negatives $C$, and the acceptable time difference between target
  syllable and correct output $\Delta t_d$, we compute the optimal
  threshold for an output element according to the definitions above:
\begin{eqnarray*}
  \textrm{true positives}_\tau &=& \textrm{size of set}_{s\in \textrm{target songs}} o_t > \tau, \left| t \leq \Delta t_d \right| \\
  \textrm{false negatives}_\tau &=& \textrm{size of set} {s\in\textrm{target songs}} - \textrm{size of set} \textrm{true positives} \\
  \textrm{false positives}_\tau &=& \textrm{size of set}_{s\in \textrm{target songs}} o_t > \tau, \left| t > \Delta t_d \right| \\
  \widehat{\tau} &=& \argmin_\tau C\textrm{false positive} + \textrm{false negatives}
\end{eqnarray*}
}

\subsubsection{Timing}

We evaluate the time taken from the presentation of each instance of
the target syllable to the firing of the detector's TTL pulse---for
example, if the target trigger point is at 200 ms with respect to the
corpus of aligned songs, and for a given presentation of the song the
detector fires at 203 ms, we take the latency on that trial to be 3
ms.

While playing the audio test file from any audio playback device, we
used the TTL output from the ground-truth channel of the audio output
as the trigger pulse for an oscilloscope, and compared it to the TTL
pulse produced by the detector, which sees only the birdsong channel
of the audio file. For this purpose we used a pulse generator (Philips
PM 5715, with a listed latency of 50 ns, jitter of $\leq 0.1\%$ or 50
ps, whichever is greater) to widen the detector's output spike to a
number much larger than the jitter ($\sim 100$ ms).  This obviates
pulse length variability in the output device by essentially
discarding the falling edge of the output pulse.  The oscilloscope is
then set to averaging mode (128-trigger average) in order to collect
timing data. The canonical signal is the trigger at $t=0$, and the
average of the detector's detection events will be seen as a
low-to-high transition with form approximating the cumulative
probability distribution function (CDF) of the detector's output in
response to the chosen song event, with individual detection events
visible as steps in this curve.

Mean latency is then given as the average time between the canonical
signal and the corresponding detection event.  It is a helpful number,
but not a critical one, since in zebra finch a detector with high but
constant latency can often be trained to trigger at a point somewhat
before the true moment of interest (for example, if triggering should
occur at the very beginning of a syllable, the detector's time window
may include previous syllable).  Often more important is the
variability in the latency?  We define jitter as the standard
deviation of the latency.

When measuring timing, it is useful to compare against a theoretical
optimal, in order to control for any imprecision in the song
extraction and alignment of our real data inherent in the method we
use (described in \cite{Poole2012}).  We propose two measures thereof:

First we test the ``ideal'' latency and jitter in the time shift used
in calculating new columns of the spectrogram. By passing a recorded
audio sequence into the detector offline, and assuming no input or
output latency, we compare how many additional audio samples beyond
the syllable are needed before the spectrogram can match the inputs
needed to trigger the neural network.  This latency reflects the FFT
size used for calculating the spectrogram, the FFT time shift between
columns in the spectrogram, and the width of the Gaussian smoothing
kernel applied to the ground truth data when training the neural
network, but ignores computation times, audio buffers, and other
operating system overhead.

Next we use a ``$\delta$-syllable'' consisting of a $\delta$ function
in the time domain, and train the network to trigger 5 ms after this
pulse.  This song is fed into the live detector.  The results for this
measurement show the latency and jitter inherent in the complete
detector including audio read and write buffers, classifier
computation, and FFT time shift aliasing, but excluding imprecision in
the song alignment as well as detection timing effects due to
differences between instances of the bird's song.

Finally, we measure latency on real bird song aligned as in
\cite{Poole2012}.  This extracts and aligns the centres of the sampled
songs, and the canonical signal is given with respect to that
alignment.  These timing measurements reflect not only detector
performance, but also the variability of the bird's song.

\subsection{Ethics}

All procedures were approved by the Institutional Animal Care and Use
Committee of Boston University (protocol number 14-029).

\section{Results}
\label{sec:results}

We have used this detector for song alignment and feedback experiments
in 5 behaving birds, and for offline analysis on a few others.  For
simplicity we present performance on 6 randomly chosen points in one
bird's song, but the results presented here are representative of
real-world performance in our lab.\marginpar{Ok?}

\fig{fig:song} shows the song for which we present accuracy and timing
results.  Six example target trigger points are shown, at 150, 200,
250, 300, 350, and 400 milliseconds after the beginning of the aligned
samples, as indicated by the red lines, and referred to henceforth as
$t^*_1\ldots t^*_6$ respectively.  The rectangles show the recognition
window (30 ms, 1--8 kHz) for each instant in the song.

The first two triggers, $t^*_1$ at 150 ms and $t^*_2$ at 200 ms, are
far from pure tones, but are closer to coloured noise.  They would be
difficult for a harmonic-stack filter bank to detect, and especially
to pinpoint in time.  $t^*_3$ (250 ms) and $t^*_6$ (400 ms) are rich
in overtones and contain downward slides with changing timbres.
$t^*_4$ (300 ms) occurs near the beginning of a more typical harmonic
stack amenable to a variety of recognition techniques, although
consistently detecting a point partway through the syllable demands a
detector that can use multiple time steps.  $t^*_5$ (350 ms) shows a
steady pitch followed by a complex changing tone structure.

An example of the detector's output for three of these syllables is
shown in \fig{fig:detection_raster}.  The total audio energy of each
song is shown as a single grayscale row.  By sorting according to time
of each syllable's detection, the rest of the song is shown in the
context of the timing for that moment.  This gives an intuition of the
timing variability within each song and across songs (which may be
responsible for some of our measured jitter).

\begin{figure}
  \includegraphics[width=\textwidth]{Fig2}
  % 9x10 inches
  % Fonts: Helvetica, scaled 130%
  \caption{Each plot shows one network output unit's responses to all
    2818 presentations of the song shown in \fig{fig:song}.  For
    simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
    $t^*_6$. The horizontal axis is time relative to the beginning of
    the aligned song, and the vertical axis is an index for the 2818
    individual song presentations. The grey shading shows the audio
    amplitude of song Y at time T. Detection events on training songs
    are shown in cyan, with detections of unseen test songs in red.  To provide an intuition of intra-song variability, songs have been
    stably sorted by the time of detection events; thus, each of the
    three detection graphs shows the songs in a different order.}
  \label{fig:detection_raster}
\end{figure}

\subsection{Accuracy}

We allowed our training software to allocate our default of 4 hidden
units per syllable, and computed a new FFT every 1.5 ms.  Because our
timing test files are designed for our stereo playback software,
allowing only one channel for the ground-truth pulse, we trained one
detector for each syllable for the following results.  In order to
eliminate the large number of variables involved in microphone, cage,
playback and re-digitising, we evaluated the neural network's accuracy
directly on the digitised recording.  When training and runtime data
are gathered on the same hardware setup, this is the digital signal
that the detector will see---only the non-song data may be different.
%Accuracy \marginpar{Should be! Verify! And/or explain that there is a bug in the LabVIEW implementation, if there in fact is. ALSO need to present a picture of the test syllable or etc\dots} is identical across the three detectors.
\noprint{
  Typical confusion matrices for these syllables are:

  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_1$
    & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.96451\% & 0.000298\%\\ 
    neg & 0.0355\% & 99.99970\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_2$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.92903\% & 0.000596\%\\ 
    neg & 0.0710\% & 99.99940\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_3$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.68062\% & 0.00194\%\\ 
    neg & 0.319\% & 99.99806\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_4$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.82257\% & 0.00239\%\\ 
    neg & 0.177\% & 99.99761\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_5$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.78708\% & 0\\ 
    neg & 0.213\% & 100\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_6$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.71611\% & 0.00552\%\\ 
    neg & 0.284\% & 99.99448\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
}

Accuracies for our six test syllables are shown in
\fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
songs were always 100\%).  For each test syllable we trained
100 different
networks, which differ from each other in the random initialisation of
network weights and on which random subset of our songs was used for
training.  Each network's accuracy is shown as a single point in each
chart in \fig{fig:accuracies}.  Some syllables are easier to identify
reliably than others, allowing detection point choice to impact the
detector's accuracy.  For difficult syllables, the training process
occasionally yields a network that performs unusually poorly, but it
is easy to spot these networks by their performance either as measured
as part of the training process or on the test audio file, and
re-train when necessary.

\begin{figure}
  % Width 8, height 3, font 120%
  \includegraphics[width=\textwidth]{Fig3}
  %\includegraphics[width=\textwidth]{accuracies-2hidden}

  \caption{Accuracy variability over 100 different training runs for each
    of the six test syllables.  Each dot shows the test-set accuracy
    for an independently trained detector.  Because the horizontal
    positions have been randomised slightly so as not to occlude same-valued
    measurements, test syllable is also indicated by colour.}
  \label{fig:accuracies}
\end{figure}

We repeated the above experiment using only 2 hidden units per syllable.  Training under this condition is much faster, but the mean true positive rate per syllable decreases by an average of 18\% across our test syllables, and the mean false positive rate increases by roughly 23\%.

\subsection{Timing}

We evaluate variability in timing performance across three variables:
FFT frame interval; syllable choice; and detector implementation.

\subsubsection{FFT frame interval}

Detector latency and jitter depend on the FFT frame rate.  Our
1.5-ms--frame default is a compromise: shorter frames increase the
precision of timing, but also increase computational requirements both
during training and at runtime.  \fig{fig:TimingVsFrame} shows how
these numbers vary over a useful range of frame rates on our ideal
detector, the Swift detector with serial output, and for the LabVIEW
detector, for both the $\delta$-syllable and for the bird song at
$t^*_4$.

\begin{figure}
  \includegraphics[width=\textwidth]{Fig4}
  \caption{Timing varies as the FFT frame interval changes.  Here we
    show results for the ideal detector and the LabVIEW and
    Swift+serial implementations, for the constructed
    $\delta$-syllable and for trigger $t^*_4$ of the bird song.  The
    lines show latency; error bars are standard deviation (jitter).  Points have
    been shifted horizontally slightly for clarity; original positions are [0.5 1
      1.5 2 4] ms.}
  \label{fig:TimingVsFrame}
\end{figure}

\subsubsection{Syllable choice}

Syllable choice impacts detector performance, but despite the variety
of syllables chosen here, performance was fairly stable across
syllables.  \fig{fig:TimingVsSyllable} shows measured timing data for
the Swift+serial detector compared to the ideal, with
$t_{\textrm{fft}}=1.5$ ms.  Given the variety of the test syllables,
the variability is surprisingly low, and is summarised here with 95\%
confidence intervals (with n=6):
\vspace{8pt}\par\noindent
\begin{tabular}{r|cc}
  Detector & Latency & Jitter \\ 
  \hline   Ideal & $-0.8\pm 0.3$ ms & $2.1\pm 0.1$ ms \\
  Swift+serial & $0.0\pm 0.6$ ms & $2.0\pm 0.1$ ms
\end{tabular}
\vspace{8pt}\par\noindent
The negative latency is due to the way in which the network responds
to the song through time: as the recognition region looks increasingly
similar to the trained match, the network's evaluation of similarity
rises, and will generally cross the triggering threshold before it
reaches its maximum value.  A heuristic as simple as triggering at an
apparent turning point after crossing the threshold might improve
timing consistency at the expense of latency, but we did not test this.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig5}
  \end{center}
  \caption{Timing data for our 6 test syllables, for the ideal and the
    Swift+serial detectors, with an FFT frame rate of 1.5 ms.  Point
    centres show latency; error bars show jitter.}
  \label{fig:TimingVsSyllable}
\end{figure}

\subsubsection{Detector Implementations}

We compared latency and jitter across our different detector
implementations for the $\delta$-syllable and $t^*_4$, again with
$t_\mathrm{fft}=1.5$ ms.  Results are as follows:
\vspace{8pt}
\begin{tabular}{l|c|c|c| c}
  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird: $t^*_4$} \\
  Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
  \hline
  Ideal & 0.66 & 0.38 & -1.0 & 2.0 \\
  LabVIEW & 1.5 & 0.69 & -0.18 & 2.4 \\
  Swift+serial & 0.88 & 1.6 & -0.075 & 2.0 \\
  Swift+audio & 5.9 & 2.6 & 1.2 & 4.9 \\
  MATLAB+serial & 8.1 & 2.3 & 3.8 & 2.0 \\
  MATLAB+audio & 13 & 5.9 & 8.4 & 5.9
\end{tabular}
\vspace{8pt}\par\noindent
These data are also plotted in \fig{fig:TimingVsDetector}, and
\fig{fig:timing} gives a more detailed view of what the timing curves
look like for the five implementations of our detector on $t^*_4$.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig6}
  \end{center}
  \caption{The different detectors for the constructed
    $\delta$-syllable and for the bird song at $t^*_4$.  Point centres
    show latency; error bars show jitter.}
  \label{fig:TimingVsDetector}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig7}
  \end{center}
  \caption{Raw timing curves for all detectors measured during
    detection of $t^*_4$ using 1.5-ms frames.  We extract the trigger events from each curve, from which we obtain the mean---latency---and
    standard deviation---jitter.}
  \label{fig:timing}
\end{figure}

\section{Discussion}
\label{sec:conclusion}

This syllable detector is appropriate for zebra finch song, and
although our tests were carried out on songs from that species, it is
also likely to work well for Bengalese finches due to their similarly stereotyped syllables.  It offers the
following benefits:
\begin{itemize}
\item The detector is accurate. False negative and false positive
  rates can be well under 0.5\% and 0.01\% respectively, and trading
  these two numbers off against each other is through a single
  relative-cost parameter.
\item Latency is generally under a millisecond, with jitter around 2
  ms.
\item Works on a wide range of target syllables using the default
  values described here, generally eliminating the need for
  hand-tuning.
\item Runs fast enough for syllable-modification experiments on
  inexpensive consumer-grade hardware, although we recommend that the
  training phase be run on a fast desktop system with 32 GB of RAM.
\item A single detector can generate different target pulses for
  multiple syllables at almost no additional computational cost during
  runtime, although training time will increase.
\end{itemize}

Although there are differences, the Swift+serial detector and the
LabVIEW implementation are roughly comparable in performance.  We
prefer the Swift implementation due to its lower hardware and software
requirements and the difficulty of debugging LabVIEW programmes.  With
serial output, MATLAB's performance is good, although its buffer
handling is sensitive to system load.

The song presented here was recorded with a fixed microphone mounted
inside the cage.  We found that higher accuracy is achieved when a
microphone is instead mounted on the bird's head, which maintains
volume and reduces changes in timbre as the bird moves.
% We have head mounted microphone data?

A common experimental paradigm requires detecting the frequency of
syllables.  Many pitch detection techniques rely on the spectrum,
which incurs no additional computational cost here since it is already
available.  For example, \cite{Canopoli2014} achieved good results
with the Harmonic Product Spectrum algorithm
\cite{Noll1970pitchdetection}.

In order to monitor syllable duration, the beginning and end of a
syllable may be detected by looking at the ratio of total energy in
the singing frequency band to the total energy, over some small time
window.  Any syllable thus identified that also contains a trigger
event may be monitored for duration.  Alternatively, the network can
be trained to recognise both the beginning and the end of the syllable
of interest.

In feedback experiments such as frequency- or duration-shifting, vocal
output changes over the course of the experiment.  The neural network
excels at identifying syllables close to its training set, so as vocal
output changes the detector may not recognise a match.  If the
detector must be robust to this shift, it may be retrained as often as
necessary as the bird learns, or data consisting of synthetically
pitch-shifted or duration-shifted target syllables over the
recognition region may be added to the training set.  We will test
these approaches in future work.

\section{Acknowledgments}
We would like to thank Winthrop F.~Gillis, William A.~Liberti, and Sanne Moorman for
their improvements to the draft.  This work was funded by NIH grants
5R01NS089679-02 and 5U01NS090454-02.  The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

\appendix

\section{Resources}
\label{sec:resources}

{\em If accepted, before final submission, all code will be moved to the Gardner Lab github repository, final release versions generated, and DOIs created as described in https://guides.github.com/activities/citable-code/ so that our results may be reproduced and the software deployed.  For now, in lieu of final release DOIs, we provide these links to our development repositories.  If you would like final versions and test datasets for review, please contact the corresponding author (bwpearre@bu.edu).}
%\marginpar{PLOS doesn't really say where such things should go.}
%\marginpar{Cleanup: mostly done.  But it would be nice to move all this to the lab's GitHub group.}
\begin{description}
  \item[Song alignment:] Jeff Markowitz's song alignment software \cite{Poole2012} can be found at
    {\tt https://github.com/jmarkow/zftftb}
\end{description}

\noindent Our implementation of the syllable
detector is available under an open-source license {\em (we are still evaluating which open source license is the most suitable for this software):}
\begin{description}
\item[Training the neural network:]\hfill
  \begin{trivlist}
  \item {\tt https://github.com/bwpearre/birds/align}
  \end{trivlist}

  \begin{description}
  \item[Installation:]\hfill
    \begin{trivlist}
    \item Clone the Github repository onto your computer.
    \end{trivlist}
  \item[Usage:]\hfill
    \begin{itemize}
      \item The main programme is {\tt learn\_detector.m}.
    \item The top section of {\tt learn\_detector.m} contains global configuration variables.  The important ones are the data directory and the bird name: as supplied, the software looks for {\tt \textless data directory\textgreater/\textless bird name\textgreater/song.mat}.  Modify as necessary.
    \item The format for the data file {\tt \textless data directory\textgreater/\textless bird name\textgreater/song.mat} is specified in {\tt birds/align/README.md}: a MATLAB file containing aligned song, nonsong (calls, cage noise, etc), and the sampling frequency.
    \item The configuration file {\tt \textless data directory\textgreater/\textless bird name\textgreater/params.m} is also described in {\tt README.md}.  If the file is empty, the programme will display an image showing the average of the song spectrograms (frequency vs.~time), tell the user how to define syllables of interest, and exit.  If {\tt times\_of\_interest} is defined, the programme will train the neural network.  {\tt params.m} can also override any of the configuration parameters at the top of {\tt learn\_detector.m}.
    \item The software will produce a neural network file, {\tt detector\_bird\_times\_\dots.mat} (the ellipses signify other information encoded in the filename).  It will also produce the audio test file described in Methods.
    \end{itemize}
  \end{description}

\item[Runtime:]\hfill
  \begin{description}
  \item[MATLAB:] {\tt https://github.com/gardner-lab/syllable-detector-matlab}
    \marginpar{Jeff: Can you flesh this out?}
    \begin{description}
    \item[Installation:]\hfill
      \begin{trivlist}
      \item Clone the Github repository onto your computer.
      \end{trivlist}
    \item[Connection:]\hfill
    \item[Usage:]\hfill
    \end{description}
  \item[Swift:] {\tt https://github.com/gardner-lab/syllable-detector-swift}
    \begin{description}
    \item[Installation:]\hfill
      \begin{trivlist}
      \item Download the most recent release from the Github repository onto your computer.
      \end{trivlist}
    \item[Connection:]\hfill
      \begin{itemize}
      \item Either connect the microphone directly into the microphone jack or through a 
      preamp into the audio input jack. If the input jack supports multiple channels (e.g., 
      audio input jacks are usually stereo), you can connect multiple audio sources and run 
      different detectors on each channel simultaneously.
      \item To generate TTL signals via the audio output, use the headphone jack. The 
      audio output must have at least the same number of channels as the input.
      \item To generate TTL signals via an Arduino, load the MATLAB arduino IO sketch 
      onto the Arduino (the sketch is available in the Github repository, and provides a 
      simple serial protocol for controlling pins). Connect the Arduino 
      to the computer via USB. The TTL signal for the first input will be available on 
      pin 7, the second input on pin 8, etc.
      \end{itemize}
    \item[Usage:]\hfill
      \begin{itemize}
      \item Use {\tt convert\_to\_text.m} to convert the Matlab output file to a format 
      easily readable by Swift. This will generate a text version of the neural network.
      \item Launch the application downloaded above. It will provide an 
      menu to select an audio input and either an audio output or an Arduino output (if 
      detected). Once you select an input and an output, a new window will show all audio 
      channels associated with the input. For each channel, select a the text-encoded version of 
      the detector created in the previous step. Once configured, press run to begin 
      monitoring the inputs.
      \end{itemize}
    \item[Customization:]\hfill
      \begin{itemize}
      \item You can clone the Github repository and use Xcode to edit the project, if you 
      are interested in extending or modifying the functionality. To hook into either 
      the raw input or output, look at the ``ViewControllerProcessor.swift'' file. The 
      {\tt Processor} class acts as a coordinator, receiving income audio, passing that
      to the detector and generating appropriate output TTL signals.
      \end{itemize}
    \end{description}
  \item[LabVIEW:] {\tt https://github.com/bwpearre/birds/align}
    \begin{description}
    \item[Installation:]\hfill
      \begin{trivlist}
      \item Clone the Github repository onto your computer.
      \end{trivlist}
    \item[Connection:]\hfill
      \begin{itemize}
      \item The microphone should be connected through a preamplifier to the data acquisition card, by default on Analog Input 0 (AI0).
      \item Outputs are on the Digital I/O (DIO) channels.  See the block diagram of {\tt nnfft.vi} for documentation, or just connect to DIO0.
      \end{itemize}
    \item[Usage:]\hfill
      Use LabVIEW to open {\tt nnfft.vi}.  A file
      chooser is provided in the front panel.  Use it to point to the network file
      produced by {\tt learn\_detector.m}, and use LabView's ``Run''
      button to begin detection.
    \end{description}
  \end{description}

\end{description}

    %\bibliography{birds}
\bibliography{birds}

\end{document}