### 2016-09-09 rebuttal.txt
Incorporate comments from Sanne on making everything more kind to the foolish reviewers.
I did not update the diff doc yet.

> Reviewer #1: The authors describe a novel technique for real-time
> identification of individual vocal gestures (syllables) in songbirds
> and provide a range of performance metrics for their new technique. I
> think that this method potentially represents an important advance in
> the execution of a very useful and widely applied method for studying
> the neural basis of vocal production and plasticity. There are however
> a few revisions that would be useful.
> 
> Also, since my comments will likely give me away anyhow, I’ll just go
> ahead and identify myself – this is Sam Sober. Nice work guys!
> 
> First, I’ll comment that a real-time detection paradigm that is fast,
> accurate, and easy to implement/modify is sorely lacking in the field,
> and that the deficiencies of current techniques (in use in my own and
> other labs) both slow down current research and likely constitute a
> barrier to entry to other labs that might otherwise make important
> contributions. So – making this method widely available and, just as
> importantly, easy to implement, would be a great step forward.
> 
> Next, here are my major comments/suggestions:
> 
> 1. Whether or not people actually use this technique is going to
> depend strongly on ease of implementation. My understanding is that
> users need not be experts in implementing MATLAB’s neural network
> toolbox (and other computational steps) to use this method, however
> the level of technical expertise required of a user is not clear in
> the manuscript, especially with respect to the various
> MATLAB/Swift/LabView realtime implementations. Specifically, although
> the computational steps are well described, it is somewhat opaque how
> much user effort is involved in setting up and running the detection
> algorithm. I suggest adding a figure/table/supplemental document that
> clarifies all required user inputs and any additional code writing
> that a user must perform. This would both constitute a compact users
> guide, outlining all steps required, and also would distinguish
> “optional” steps (resetting number of hidden units per syllable, etc.)
> where users could use default values, from “required” steps (choosing
> detection points, etc.).

We try to provide a complete description of the method, but since we
are also distributing our implementation, it seems like a great idea
to do as you suggest.  Therefore we have added
installation/configuration instructions in Appendix A.

> 2. The discussion of latency/timing is somewhat confusing, and I
> believe it omits what I (and other investigators) consider to be the
> most crucial metric of detection performance – the variation in
> latency from the onset of the targeted syllable to the firing of the
> TTL pulse.

We had extensively characterised the timing from the completion of the
target recognition region to the firing of the TTL pulse.  We have
tried to clean up those references and clarify our description: if the
detection point is at the onset of a syllable, then our results apply
(see the paragraph at line 361 in the manuscript, or the paragraph at
line 375 in the diff document).

> Charlesworth et al. (“Learning the microstructure of successful
> behavior”) showed that learning after white noise targeting is
> temporally localized at less than 10 msec timescale. Furthermore, in
> my own and other labs, the key predictor of whether online targeting
> will successfully drive learning is the standard deviation of
> detections/triggers relative to syllable onset, with syllable onset
> defined by some measure of sound amplitude for each individual
> iteration of the syllable. I believe (and please correct me if I’m
> wrong) that the song alignment technique used here does not provide
> this statistic on a syllable-by-syllable basis.

((1a)) That is correct, but the alignment does appear to be
sufficiently accurate:

We refer to this syllable-by-syllable adjustment to the alignment as
micro-realignment.  Even without that micro-realignment, jitter was
generally around 2 ms, as can be seen throughout Section 3.2,
especially Figs. 4-7.  Since this is well within the 10-ms window
suggested by the literature, we decided that the improvement that
could be achieved by micro-realigning song was unimportant.  Indeed,
our measured jitter is barely above the theoretically optimal jitter
given our detection parameters, so the average micro-realignment error
cannot be significant.

> So – please provide this measure where appropriate in Results. I
> suspect that the jitter measurements for real-time performance will
> not be much different with this metric than what is currently
> reported, but it would be very useful nonetheless.

We believe that we have provided a sufficient analysis of jitter, and
that, while the improvement you suggest could only decrease our jitter
measurement, the performance without that enhancement is good enough
that the improvement we might see is not very important.

> 3. The authors should quantify additional examples of real-time
> performance. Most analysis focuses on the song of a single bird, and
> real-time latency performance is only reported for detection point
> t4 (which, given the strong harmonic content of t4, I suspect is the
> easiest syllable to target). The authors should provide similar
> performance metrics for the other detection times.

Actually, we had already done so: see Fig. 5 (and section 3.2,
"syllable choice").

> Also, the authors should provide performance metrics for at least
> one additional bird;

((1b)) We have achieved similar results on quite a few birds, and
present only one bird simply for reasons of conciseness.  Since the 6
target points in that bird's song are essentially random (50-ms
intervals in a song that includes a range of typical sound types and
does not have any 50-ms periodicity), we just added a comment that
these results are indeed representative (e.g. the beginning of Section
2.3).

> although the three syllables shown in Figure 1 look like fairly
> typical ZF syllables, a method such as this one should be shown to
> work on multiple syllables/animals. This additional bird could be the
> zebra finch shown in Fig 8-9. Alternately, given that many researchers
> use Bengalese finches in real-time detection experiments,
> demonstrating the efficacy of the method in a Bengalese would be
> useful.

It would.  But we don't currently have data sets for other species,
and we feel that it is more important to get the technique out there
quickly than to continue to characterise it in more detail.

> 4. I am ambivalent about the use of the fiber photometry data in this
> paper. Although this looks like a great technical achievement, I see
> two problems with including it here. First is that it is not actually
> necessary to demonstrate that the targeting software works – this
> could be accomplished just by quantifying the accuracy of the white
> noise hits shown in Figure 8, or by showing that white noise targeting
> using the new method successfully drives learning (indeed, driving
> learning is what most potential users of this method will likely care
> about). Second is that for the photometry data to be presented, more
> analysis is needed. How precisely was the GCaMP6s/f confined to HVC
> shelf versus HVC or more ventral areas? If there was spillover, how
> would this affect interpretation of the data shown in Fig. 9? If
> individual animals were co-infected with GCaMP6s and GCaMP6f, which
> expressed more robustly and/or contributed more to the signal?
>
> In general, it seems odd to use photometry – itself a methodological
> innovation in songbirds – to test the implementation of another
> innovative method. I think the authors should either (1) provide a
> more detailed analysis of the photometry technique, (2) confine
> analysis to real-time implementation of white noise (WN) targeting,
> without photometery data, or (3) demonstrate that pitch-contingent WN
> drives vocal learning with the new software. Among these options I
> think (3) would be the most useful, but I leave this up to the Editor
> and the authors.

((1c)) These are very good questions.  We wanted to throw in a
real-world test case, but you are right that it is not the most
appropriate one, and our other experimental test cases are far from
publication as yet.  So we've removed this section.

> Minor comments:
> 
> Line 318-319 – As I mentioned above, I find the discussion of
> timing/latency a bit confusing. For example – “time taken from the
> presentation of the target syllable to the firing of the TTL pulse” –
> I can’t tell whether the first time point is the onset time of the
> target syllable on a trial-by-trial basis or something averaged across
> songs with variable timing.

See my (1a) above.  I hope we've clarified this.

> Figure 3 – Although the “correct detection” performance is great for
> all syllables, I notice that it’s a bit worse for t3 and t6. Is this
> because these two syllables, which are acoustically very similar, are
> mistaken for each other? I’m guessing that the LabView code that my
> lab currently uses would never be able to distinguish between these,
> so the fact that performance sinks to “only” ~99% for these is a major
> strength of the technique, and you might want to point this out.

Yes, the software sometimes confuses between these two.  Noted.

> Please clarify what data/parameters are saved by the current version
> of the software. Most users will want the software to save detailed
> records of acoustic input, trigger/detection times, time-dependent
> evaluation of similarity, etc. for each song trial. If no such data
> are saved, it would be useful to indicate (either in the manuscript,
> in the code, or both) how this could be added.

None--we use other software for that.  But the new user's guide
(Appendix A) now includes a pointer for where to go to make such
modifications for the Swift implementation.

> Reviewer #2: The authors have performed a useful study to examine the
> suitability of diverse hardware platforms for real time vocal syllable
> detection. Many solutions to this problem exist already but the
> current study seemingly outperforms these. Making the software open
> access will be particularly useful for researchers in the field.
> 
> The strength of the study is that their solution achieves low latency
> and low jitter. For detecting target points and delivering TTL Pulses
> during song they implemented a neural network detector on 3 platforms:
> LabView (hard to debug), MATLAB (slowest), and Swift+Serial (fast, MAC
> based).  Their approach for syllable detection is a neural network
> (with 2 layers) that requires relatively little computation
> time. Overall the work would be outstanding if the authors had also
> evaluated an FPGA solution, but presumably such an approach is highly
> demanding on the programming side.
> 
> The main concern with the presentation of the work is that the
> classification could be better evaluated. First of all, they only
> train and test on six target points of only one song.

We discuss one song in the paper, mostly for convenience, but we've
added results are representative of what we see on many birds.  See
(1b) above.

> Most importantly, they do not do a life bird test.

I'm not sure what this means.  The fibre photometry test is a live
bird test (and ongoing research in our lab uses the detector with
great success), but all quantitative tests must be done on recordings
in order to systematically analyse accuracies of various detectors.

> In fact, they haven't included calls, wing flaps, and other noises
> in their training or testing sets.  Thus, it seems there is no
> realistic evaluation of their system.

That is an excellent point!  We have remedied this.  See lines 94, 163
and 322 in the diff document.  Also, the new Results/Accuracy graph
(Fig. 3) shows accuracy figures including lots of cage noise samples
(about half an hour's worth, as now described).  Correct detections
and incorrect detections have both somewhat decreased from our earlier
results.

> Minor:
> 
> Their method requires to find an "optimal threshold" for the output of
> the network but they don't say how they choose it. Perhaps cross
> validation could help (their network does regression, but of course
> the purpose is to output a binary signal, so the evaluation needs to
> be binary).

We do say how we choose it: see Sec. 2.1, "Computing optimal output
thresholds".

> The paper has quite a few sections on fiber photometry, but I find
> these section quite irrelevant to the main point of the paper, which
> is to detect syllables in an audio recording. Perhaps the photometry
> could be left out?

We've removed it.  See (1c) above.

> How robust is the method to errors in the training set?

There is an infinite space of possible errors in the training set, so
this question is impossible to answer well.  However, no effort was
made to compensate for errors in the alignment tool, and the results
aren't bad, so I guess it depends on what kinds of errors you're
worried about.

> the writing and presentation could be improved.
> e.g. fig 1 increase font size of labels

Thanks.  We've increased the font size and improved the writing.

> Line 189: It takes so long to train a network with just 2 to 4 hidden
> neurons?

Big data sets, and we made little effort to optimise offline
(training) speed.  But we changed the training algorithm and now
things are about 10 times faster.

> maybe reducing the training data would help (reduce the
> negative targets only).

We preferred your other suggestion--that we increase the negative
targets (adding cage noise, etc).  Training speed shouldn't matter
much compared to online performance, but the change to scaled
conjugate gradient training has sped things up considerably.

> Line 203: How was Cn chosen?

Arbitrarily.  Now noted in the text.  Thanks.

> Line 201: gradient descent, not hillclimbing (on cost). cross
> validation.

My comment applied to either one, but that text, which just described
something we didn't do, was unnecessary, so I removed it.

> Line 292: needs real data including all kinds of sounds present in the
> cage

Done.

> Line 428: robustness to these variables needs to be documented.

We documented robustness to some choices.  There are always more, but
this was supposed to be a 5-page paper.  We have to stop somewhere.

> Line 238: what is no user loads?

No loads from users (i.e. operating system processes only).  Clarified
in the text.

> Fig. 8, what are the dark lines intermingled between the yellow lines
> at the bottom quarter of the figure? false negatives?

Removed as per (1c).

> Reviewer #3: This paper describes a program to detect syllable in ZF
> songs directed at the ZF brain and learning community. The main method
> relies on neural network learning on the zscored spectrogram and
> optimization techniques to diminish latency and detection time
> variations.
> 
> I cannot recommend the paper for publication for the following reasons:
> 
> - there does not seem to be any real and documented innovation here to
>   justify a publication - it seems more like several enhancements
>   designed for a specific use in a neuroscience lab

Although there is no machine learning innovation, we believe that the
combination of making available a useful tool and fairly thoroughly
documenting its performance constitute a significant contribution to
the field.

> - which lead to my second reason: the tool and the data required is
>   too narrowly focused on one particular study on ZF. My contention is
>   that this would be usable for a very limited number of teams.

We agree with you, although several have noted that this model is
extremely powerful and the absence of readily available tools such as
this one is a barrier to entry into precisely timed zebra finch
experiments.

> Several points:
> 
> a) From a machine learning community point of view the article is
> extremely frustrating ; nothing is done in the statistics of syllables
> to justify the architecture and its hyper parameters.

Agreed.  However, it is not a machine learning paper, but rather a
paper that happens to use a simple machine learning tool to build
something useful.

> b) also the dataset is super standardized and it should have been
> tested on others labs data and why not different 'strains' of ZF
> (American, European and Wild type).

Yes, more birds/species would always be better, but see (1b).  The
characteristic variation in song between different 'strains' appears
to be much less than the variation in song between different birds
within the one strain that we have tested on, or even within the
different moments in the song of one bird that we have used as our
example.

> In other word I highly doubt that the authors' system can be
> generalised (even though they claim it can be extended to another
> species - without providing any insights on why this would be true).

Some insight added, but we agree that it's a tool for a specific
community.  As is any tool.

> c) the training is done on around 3k song annotated and aligned and
> never provide the readers and potential user how to obtain such
> dataset.

This is incorrect.  We cited the article in which alignment software
is described, and we linked to the alignment software.  The only step
that we have not described is using a recording device, which is
outside the scope of the paper.  However, if you are curious, there is
a nice tutorial at
http://soundanalysispro.com/manual-1/chapter-2-installation-of-sap2/hardware-installation-building-training-boxes

> With my own experience on working with birds in social context, such
> dataset is extremely difficult to obtain without very invasive and
> constrained method.  In the present manuscript, it has been
> obtained, presumably, by letting a bird to its own device in a small
> cage for hours to extract songs. Many experiments do not want such
> harsh treatment.

While learning to detect syllables of specific birds using samples
obtained in the wild would be an interesting research problem, it is
outside the scope of the current paper.  Instead, we only claim to
address a need encountered in a great deal of zebra finch research,
using common protocols.

> d) Also here to digress a bit, it is not specified if these are
> directed and undirected songs (or both). The distinction can be of
> import because they have different syllable statistics and of course
> this should import on the training and/or the prediction.

Good point.  Now specified (undirected) in Section 2.1.

> e) It is also assume that the user has means to extract / align song
> automatically event though this is never mentioned in the
> manuscript.

Is this the same point as your (c)?

> f) the experiment with Fibre Photometry is really nice but is a
> distraction from the main paper and brings nothing to the story

We've removed it (see (1c)).

> g) very minor: its Levenberg (not Levenburg) and it is a very unusual
> choice here (usually some kind of enhanced backpropagation are used)
> probably related to the black-boxy nature of the matlab toolbox

Thanks!  It's a good point.  Levenberg-Marquardt is the default for
MATLAB's toolbox, and we wanted to use as many defaults as we could
(preserving the black-boxy nature of third-party tools as much as
possible in order to emphasise that the specific choice probably
doesn't really matter very much--hopefully any regression tool would
work roughly similarly!).  But it did turn out in the end that Scaled
Conjugate Gradient produced similar results faster.

> in short, all these criticisms in fact converge to the main
> weakness: this paper/software is directed to a very specific
> community both in scope and type of data needed for it to work and
> the authors should aim to a journal directed at this community.

We believe that the paper presents a valuable technique and a far more
thorough characterisation of finch syllable detection accuracy and
timing than has previously been available.  It is our hope that the
performance, ease of use, simplicity, and low hardware cost of this
method will speed the development of a small but growing field.